---
title: "Case-Study-1-Beer_Breweries"
author: "David Wei, Neil Benson"
date: "6/15/2020"
output: html_document
---

## Adding In Files
_Import File & General Libraries_
## this 
```{r}
library(ggplot2) 
library(ggthemes)
library(dplyr) 
library(tidyverse)
library(forcats)

beer <- read.csv("C:/Users/David/Google Drive/Masters/Data Files/Beers.csv")
brew <- read.csv("C:/Users/David/Google Drive/Masters/Data Files/Breweries.csv")
#beer <- read.csv("C:/Git/MSDS_6306_Doing-Data-Science/Unit 8 and 9 Case Study 1/Beers.csv")
#brew <- read.csv("C:/Git/MSDS_6306_Doing-Data-Science/Unit 8 and 9 Case Study 1/Breweries.csv")

head(beer,5)
head(brew, 5)
```

## Once the files are merged in, 
_Data Tidying_
```{r}
# left join brew into beer based on brew_id, then renaming and sorting columns
beer_df = merge(x=beer, y=brew, by.x="Brewery_id", by.y="Brew_ID", all.x=TRUE) %>% 
  rename("Oz"=Ounces, "Brewery_ID"=Brewery_id, "Beer_Name"=Name.x, "Brewery_Name" =Name.y)
beer_df <- beer_df[c(1,8,3,2,9,10,4,5,6,7)]
# trimming leading and tailing spaces
library(stringr)
beer_df$State <- str_trim(beer_df$State,side="both")
head(beer_df, 5)

# checking for empty data and imputating NA values
library(naniar)
library(corrplot)
library(imputeTS)
vis_miss(beer_df) + xlab("Data Columns")

# imputating NA values with the mean
beer_df_cleaned <- na_mean(beer_df, option = "mean")
head(beer_df_cleaned, 5)
```

_Analysis #1: How many breweries are present in each state?_
```{r}
# beer_df_cleaned %>% group_by(State, Brewery_ID) %>% tally summarise(brew_count_per_state = count())
# beer_df_cleaned %>% arrange(desc(Max_IBU)) %>% ggplot(aes(x = State,fill=State)) + geom_histogram(stat="count") + theme(legend.position = "none") + ggtitle("# of Breweries per State")

#beer_df_cleaned %>% group_by(State,Brewery_ID) %>% ggplot(aes(x = State,fill=State)) + geom_bar(stat="count") + theme(legend.position = "none") + ggtitle("# of Breweries per State")
beer_df_cleaned %>%
  group_by(State) %>%
  summarise(distinct_breweries = n_distinct(Brewery_ID)) %>% 
  mutate(State = fct_reorder(State, desc(distinct_breweries))) %>%
  ggplot(aes(x = State, y = distinct_breweries, fill=State)) + 
  geom_bar(stat="identity") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8)) + 
  ggtitle("# of Breweries per State")

```
_Analysis #2: Merge Beer and Breweries, print first and last 6 obs_
```{r}
head(beer_df_cleaned, 6)
tail(beer_df_cleaned, 6)
```
_Analysis #3: Addressing missing values in each column_
```{r}
vis_miss(beer_df) + xlab("Data Columns")
```
_Analysis #4: Compute Median ABV and IBU for each state and plot_
```{r}
# creating logic column as calculation of medians and storing each as seperate dfs
ABV_summary <- beer_df_cleaned %>% group_by(State) %>% summarise(Median_ABV=median(ABV), Max_ABV=max(ABV)) %>% as.data.frame()
IBU_summary <- beer_df_cleaned %>% group_by(State) %>% summarise(Median_IBU=median(IBU), Max_IBU=max(IBU)) %>% as.data.frame()
# joining the 2 dfs
summary_df <- merge(x=ABV_summary, y=IBU_summary, by="State", all=TRUE)
# scaling ABV and IBU for easier comparison

summary_df$median_ABV.scaled = scale(summary_df$Median_ABV)
summary_df$median_IBU.scaled = scale(summary_df$Median_IBU)
summary_df$max_ABV.scaled= scale(summary_df$Max_ABV)
summary_df$max_IBU.scaled = scale(summary_df$Max_IBU)
head(summary_df, 5)
#comparing median ABV vs IBU per state
library(reshape)
summary_df_pivot <- melt(summary_df)

summary_df_pivot %>% filter(variable=="median_ABV.scaled" | variable=="median_IBU.scaled") %>% ggplot(aes(State,value,fill=variable)) + theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8)) + 
  geom_bar(stat="identity", position="dodge") + labs(y="Scale", title="Comparison of Median ABV vs IBU per State")

# comparing median ABV per state
summary_df %>% arrange(desc(Median_ABV)) %>% mutate(State=fct_reorder(State, Median_ABV)) %>% 
  ggplot(aes(State, Median_ABV,fill=State)) + geom_bar(stat="identity", position="dodge", fill="#F8766D") + theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8)) + 
  scale_y_continuous("Med ABV (%)",label=scales::percent)  + labs(title="Median ABV per State")


# comparing median IBU per state
summary_df %>% arrange(desc(Median_IBU)) %>% mutate(State=fct_reorder(State, Median_IBU)) %>%
  ggplot(aes(State, Median_IBU,fill=State)) + geom_bar(stat="identity", position="dodge",fill="#00BFC4") + theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8))+ labs(title="Median IBU per State")

```
_Analysis #5: Which state has maximum ABV beer? Which state has most bitter(max IBU) beer?_
```{r}
# comparing ABV vs IBU max

# myColors <- c("#7CAE00", "#C77CFF")
myColors <- c("#56ddc5", "#4699dd")
summary_df_pivot %>% filter(variable=="max_ABV.scaled" | variable=="max_IBU.scaled") %>% ggplot(aes(State,value,fill=variable)) + theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8)) + 
  geom_bar(stat="identity", position="dodge") + scale_fill_manual(values=myColors) + labs(y="Scale", title="Comparison of Max ABV vs IBU per State")

# comparing ABV per state
summary_df %>% arrange(desc(Max_ABV)) %>% mutate(State=fct_reorder(State, Max_ABV)) %>% slice(1:10) %>% 
  ggplot(aes(State, Max_ABV,fill=State)) + geom_bar(stat="identity", position="dodge",fill="#56ddc5") + theme(legend.position="none") + 
  scale_y_continuous("Max ABV (%)",label=scales::percent) + labs(title="Top 10 States with Most Alcoholic Beers")

# comparing IBU per state
summary_df %>% arrange(desc(Max_IBU)) %>% mutate(State=fct_reorder(State, Max_IBU)) %>% slice(1:10) %>%
  ggplot(aes(State, Max_IBU,fill=State)) + geom_bar(stat="identity", position="dodge",fill="#4699dd") + 
  ylab("Max IBU") + theme(legend.position="none") + labs(title="Top 10 States with Most Bitter Beers")


# validating accuracy of data
summary_df %>% filter(State=="CO")
summary_df %>% filter(State=="OR")
```

_Analysis #6: Commenting on summary stats and distributino of ABV variable._
```{r}
# refer to Analysis #5 outputs
summary(beer_df_cleaned$ABV)
hist(beer_df_cleaned$ABV, breaks=10, border="black", col="#7CAE00", xlab="ABV (%)", main="Beer ABV Distribution")
summary_ABV <- summary_table(beer_df_cleaned, )

summary_ABV <- beer_df_cleaned %>% ggplot(aes(beer_df_cleaned$ABV,fill=State)) + geom_density()
summary_ABV

library(psych)
describe(beer_df_cleaned$ABV)

```
_Analysis #7: Relationship between IBU and ABV?_
```{r}
# logging variables to see if it helps improve correlation
beer_df_cleaned$ABV.log <- log(beer_df_cleaned$ABV)
beer_df_cleaned$IBU.log <- log(beer_df_cleaned$IBU)

# install.packages("ggExtra")
library(ggExtra)

ABV_IBU_rel <- beer_df_cleaned %>% group_by(State) %>% ggplot(aes(x=ABV, y=IBU)) + geom_point() + geom_smooth(method="lm") +
  scale_x_continuous("Max ABV (%)",label=scales::percent) + 
  labs(title="Relationship between ABV and IBU",
       caption="Source: beer, brewery data")
ABV_IBU_rel_hist <- ggMarginal(ABV_IBU_rel, type="histogram", fill="slateblue")#, xparams=list(bins=10), yparams=list(bins=10))
ABV_IBU_rel_hist

cor(beer_df_cleaned$ABV, beer_df_cleaned$IBU)
cor(beer_df_cleaned$ABV.log, beer_df_cleaned$IBU.log)
cor(beer_df_cleaned$ABV, beer_df_cleaned$IBU.log)
cor(beer_df_cleaned$ABV.log, beer_df_cleaned$IBU)
```
_Analysis #8: Difference of IBU and ABV between IPAs and other Ales?_
```{r}
# goal here is to use kNN to classify if a beer is Ale/IPA/Other based on it's ABV and IBU
# viewing all types of beer styles
beer_df_cleaned %>% group_by(Style) %>% select(Style)

# converting existing beers into type categories
beer_df_cleaned$Style[grep("Ale", beer_df_cleaned$Style)]
beer_df_cleaned$Style[grep("IPA", beer_df_cleaned$Style)]
beer_df_cleaned$Style[grep("Lager", beer_df_cleaned$Style)]

# creating string match pattern
beer_df_cleaned$Beer_Category <- ifelse(grepl("Ale", beer_df_cleaned$Style), "Ale", 
                                        ifelse(grepl("IPA", beer_df_cleaned$Style) | grepl("India Pale Ales", beer_df_cleaned$Style), "IPA", "Other"))
head(beer_df_cleaned, 50)

# kNN
library(class) # for knn
library(caret) # for confusionMatrix
match("ABV", names(beer_df_cleaned)) # column 7
match("IBU", names(beer_df_cleaned)) # column 8
# running kNN k=5

knn.cv(beer_df_cleaned[,c(7,8)],beer_df_cleaned$Beer_Category,prob = TRUE, k = 5)
beer_df_cleaned$pred_beer_cat <- knn.cv(beer_df_cleaned[,c(7,8)],beer_df_cleaned$Beer_Category,prob = TRUE, k = 5)
confusionMatrix(table(beer_df_cleaned$pred_beer_cat,beer_df_cleaned$Beer_Category))
```

_Analysis #9: EDA inference_
```{r}
head(beer_df_cleaned, 5)
```
