---
title: "Case Study 1 - Beer and Breweries"
author: "By David Wei, Neil Benson"
date: "06/15/2020"
output: html_document
---

## Import File & General Libraries
_We first start with importing the main libraries and then reading in files and showing first 5 lines of both beer and brewery datasets._
```{r,warning=FALSE,message=FALSE}
library(ggplot2) 
library(ggthemes)
library(dplyr) 
library(tidyverse)
library(forcats)


#beer <- read.csv("C:/Users/David/Google Drive/Masters/Data Files/Beers.csv")
#brew <- read.csv("C:/Users/David/Google Drive/Masters/Data Files/Breweries.csv")
beer <- read.csv("C:/Git/MSDS_6306_Doing-Data-Science/Unit 8 and 9 Case Study 1/Beers.csv")
brew <- read.csv("C:/Git/MSDS_6306_Doing-Data-Science/Unit 8 and 9 Case Study 1/Breweries.csv")

```
#### Reviewing the first 6 records of beer
|Name 				          |Beer_ID   		  |ABV 		    |IBU 	  |Brewery_id 	  |Style 							            |Ounces	|
|-----------------------|---------------|-----------|-------|---------------|-------------------------------|-------|
|Pub Beer    			      |1436 			    |0.050  	  |NA     |409			      |American Pale Lager     		    |12		  |
|Devil's Cup    		    |2265 			    |0.066  	  |NA     |178			      |American Pale Ale (APA) 		    |12		  |
|Rise of the Phoenix    |2264 			    |0.071  	  |NA     |178			      |American IPA     				      |12		  |
|Sinister    			      |2263 			    |0.090  	  |NA     |178			      |American Double / Imperial IPA |12		  |
|Sex and Candy    		  |2262 			    |0.075  	  |NA     |178			      |American IPA     				      |12		  |
|Black Exodus    		    |2261 			    |0.077  	  |NA     |178			      |Oatmeal Stout     				      |12    	|  

#### Reviewing the first 6 records of brewery
|Brew_ID|Name					             |City          |State |
|-------|--------------------------|--------------|------|
|1		  |NorthGate Brewing 		     |Minneapolis   |MN    |
|2		  |Against the Grain Brewery |Louisville    |KY    |
|3		  |Jack's Abby Craft Lagers  |Framingham    |MA    |
|4		  |Mike Hess Brewing Company |San Diego     |CA    |
|5		  |Fort Point Beer Company   |San Francisco |CA    |
|6		  |COAST Brewing Company	   |Charleston    |SC    |     


## Analysis #1
_Once our data has been imported we can now begin our analysis starting with our first question of determing how many breweries are 
present in each state. After counting the number of unique breweries per state, we found that Colorado leads in having 47 breweries._
```{r,warning=FALSE,message=FALSE}
brew %>%
  group_by(State) %>%
  summarise(Breweries = n_distinct(Brew_ID)) %>% 
  mutate(State = fct_reorder(State, desc(Breweries))) %>%
  ggplot(aes(x = State, y = Breweries, fill=State)) + 
  geom_bar(stat="identity") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8)) + 
  ggtitle("# of Breweries per State")
```


## Analysis #2
_After the beer and brewery files have been loaded in as dataframes, we then joined, renamed and re-arranged the 2 datasets 
together to create a single tidy source of data for ease of use. Output of final dataset provided below._
```{r,warning=FALSE,message=FALSE}
library(stringr)

## Data Tidying


# left join brew into beer based on brew_id, then renaming and ordering columns
beer_df <- merge(x=beer, y=brew, by.x="Brewery_id", by.y="Brew_ID", all.x=TRUE) %>% 
  plyr::rename(c("Ounces"="Oz", "Brewery_id"="Brewery_ID", "Name.x"="Beer_Name","Name.y"="Brewery_Name"))
 

# reording the columns
beer_df <- beer_df[c(1,8,3,2,9,10,4,5,6,7)]


# trimming leading and tailing spaces
beer_df$State <- str_trim(beer_df$State,side="both")

```

#### Reviewing the first 6 records after join
|Brewery_ID |Brewery_Name 					         |Beer_ID|Beer_Name    				       |City		 	     |State	 |ABV 	 |IBU	   |Style 									             |Oz|
|-----------|--------------------------------|-------|---------------------------|---------------|-------|-------|-------|-------------------------------------|--|
|1 			    |NorthGate Brewing  			       |2692   |Get Together 				       |Minneapolis	   |MN 	   |0.045  |50		 |American IPA 							           |16|
|1 			    |NorthGate Brewing  			       |2691   |Maggie's Leap				       |Minneapolis 	 |MN 	   |0.049  |26     |Milk / Sweet Stout 					         |16|
|1 			    |NorthGate Brewing  			       |2690   |Wall's End 				         |Minneapolis    |MN 	   |0.048  |19     |English Brown Ale 						       |16|
|1 			    |NorthGate Brewing  			       |2689   |Pumpion 					         |Minneapolis    |MN 	   |0.060  |38     |Pumpkin Ale 							           |16|
|1 			    |NorthGate Brewing  			       |2688   |Stronghold 				         |Minneapolis    |MN 	   |0.060  |25     |American Porter 						         |16|
|1 			    |NorthGate Brewing  			       |2687   |Parapet ESB 				       |Minneapolis    |MN 	   |0.056  |47 	   |Extra Special / Strong Bitter (ESB)  |16|

#### Reviewing the last 6 records after join
|Brewery_ID |Brewery_Name 					         |Beer_ID|Beer_Name          		     |City 			     |State  |ABV 	 |IBU    |Style 									             |Oz|
|-----------|--------------------------------|-------|---------------------------|---------------|-------|-------|-------|-------------------------------------|--|
|556        |Ukiah Brewing Company      	   |98     |Pilsner Ukiah         		 |Ukiah    		   |CA 	   |0.055  |NA     |German Pilsener 						         |12|
|557        |Butternuts Beer and Ale      	 |52  	 |Heinnieweisse Weissebier 	 |Garrattsville  |NY 	   |0.049  |NA     |Hefeweizen 							             |12|
|557        |Butternuts Beer and Ale      	 |51     |Snapperhead IPA 			     |Garrattsville  |NY 	   |0.068  |NA     |American IPA 							           |12|
|557        |Butternuts Beer and Ale      	 |50     |Moo Thunder Stout 			   |Garrattsville  |NY 	   |0.049  |NA     |Milk / Sweet Stout 					         |12|
|557        |Butternuts Beer and Ale      	 |49     |Porkslap Pale Ale 			   |Garrattsville  |NY 	   |0.043  |NA 	   |American Pale Ale (APA) 				     |12|
|558 		    |Sleeping Lady Brewing Company   |30 	   |Urban Wilderness Pale Ale  |Anchorage    	 |AK 	   |0.049  |NA     |English Pale Ale 						         |12|


## Analysis #3
_Addressing missing values in each column, we found that almost half (~42%) of IBU data was missing, and a little over 2.5% of AV data was missing. 
Imputation was applied to correct the missing data, however, we found that impuation actually underfitted the data values for IBU and ABV
and so we added an additional layer of logic to the imputation by applying means of IBU and ABV per style of beer to provide higher accuracy._
```{r,warning=FALSE,message=FALSE}
library(naniar)
library(corrplot)
library(imputeTS)

# checking for empty data
vis_miss(beer_df) + xlab("Data Columns")


# get all non NA for ABV and IBU to calculate mean
abv = beer_df[!is.na(beer_df$ABV),]
ibu = beer_df[!is.na(beer_df$IBU),]


# get mean ABV and IBU by style
beer_mean_ABV <<- abv %>% 
  group_by(Style) %>% 
  summarize(mean_ABV = round(mean(ABV),3))

beer_mean_IBU <<- ibu %>% 
  group_by(Style) %>% 
  summarize(mean_IBU = round(mean(IBU),0))


# fill in missing values for IBU where style has no recorded IBU
Style <- c('American Malt Liquor','Braggot','Cider','Flanders Red Ale','Kristalweizen','Low Alcohol Beer','Mead','Rauchbier','Shandy')
mean_IBU <- c(20, 30, 0, 20, 12.5, 15, 1.66, 25, 10.208)
IBU_missing <- data.frame(Style, mean_IBU)
beer_mean_IBU <- rbind(beer_mean_IBU,IBU_missing)


# left join IBU and ABV mean values by style
mean_abv_ibu <- merge(x=beer_mean_ABV, y=beer_mean_IBU, by.x="Style", by.y="Style", all.x=TRUE)


# function to impute mean for ABV and IBU based on means of style
naMeanFill <- function (Style, Column, Colname) {
  if ( Colname == 'ABV') 
    { 
      mean = mean_abv_ibu$mean_ABV[ mean_abv_ibu$Style == Style]
    } 
  else
    { 
      mean = mean_abv_ibu$mean_IBU[ mean_abv_ibu$Style == Style]
    }
  Column[is.na(Column)] <- mean
  return(Column)  
}


# vectorize the function
naMeanFill_v <- Vectorize(naMeanFill)


# fill-in NAs for ABV and IBU
beer_df$ABV <-naMeanFill_v(beer_df$Style,beer_df$ABV,'ABV')
beer_df$IBU <-naMeanFill_v(beer_df$Style,beer_df$IBU,'IBU')
beer_df$IBU <- as.integer(beer_df$IBU)


# imputating NA values with the mean
beer_df_cleaned <- beer_df

vis_miss(beer_df_cleaned) + xlab("Data Columns")
```

#### Address the missing values in each column
IBU missing 41.7% of data
ABV missing 2.57% of data
All other data noted as complete
Missing data was filled in using the average ABV and average IBU by beer style


## Analysis #4
_The Median alcohol content (ABV) and international bitterness unit (IBU) for each state was looked into by scaling the ABU and IBU coefficients 
at a comparitive level due to how different each unit of measurement is. Additionally, we individually plotted the mean measurements per state and 
found that most beers fell within relative range of each other with only a handful of states significantly differing from the rest._
```{r,warning=FALSE,message=FALSE}

# creating logic column as calculation of medians and storing each as seperate dfs
ABV_summary <- beer_df_cleaned %>% group_by(State) %>% summarise(Median_ABV=median(ABV), Max_ABV=max(ABV), Mean_ABV=mean(ABV)) %>% as.data.frame()
IBU_summary <- beer_df_cleaned %>% group_by(State) %>% summarise(Median_IBU=median(IBU), Max_IBU=max(IBU), Mean_IBU=mean(IBU)) %>% as.data.frame()


# joining the 2 dfs
summary_df <- merge(x=ABV_summary, y=IBU_summary, by="State", all=TRUE)

# scaling ABV and IBU for easier comparison
summary_df$median_ABV.scaled = scale(summary_df$Median_ABV)
summary_df$median_IBU.scaled = scale(summary_df$Median_IBU)
summary_df$max_ABV.scaled= scale(summary_df$Max_ABV)
summary_df$max_IBU.scaled = scale(summary_df$Max_IBU)
```

#### Reviewing summary_df
|State|Median_ABV|Max_ABV|Median_IBU|Max_IBU| median_ABV.scaled|median_IBU.scaled| max_ABV.scaled| max_IBU.scaled|
|-----|----------|-------|----------|-------|------------------|-----------------|---------------|---------------|
|AK   |0.056     |0.068  |34.0      |71     |0.006867089       |0.02227067       |-1.4023857     |-1.1581394     |
|AL   |0.060     |0.093  |39.5      |103    |0.940791179       |0.71637307       |0.2702446      |0.4235204      |
|AR   |0.052     |0.061  |36.0      |46     |-0.927057001      |0.27467154       |-1.8707222     |-2.3938111     |
|AZ   |0.057     |0.095  |22.0      |99     |0.240348111       |0.77947329       |0.6716758      |1.0166429      |
|CA   |0.058     |0.099  |40.0      |115    |0.473829134       |-1.49213459      |0.4040550      |0.2258130      |


```{r,warning=FALSE,message=FALSE}
#comparing median ABV vs IBU per state
library(reshape)
summary_df_pivot <- melt(summary_df)

summary_df_pivot %>% filter(variable=="median_ABV.scaled" | variable=="median_IBU.scaled") %>%
  # mutate(State = fct_reorder(State, desc(distinct_breweries))) %>%
  ggplot(aes(State,value,fill=variable)) + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8)) + 
  geom_bar(stat="identity", position="dodge") + 
  labs(y="Scale", title="Comparison of Median ABV vs IBU per State")

# comparing median ABV per state
summary_df %>% arrange(desc(Median_ABV)) %>% 
  mutate(State=fct_reorder(State, Median_ABV)) %>% 
  ggplot(aes(State, Median_ABV,fill=State)) + 
  geom_bar(stat="identity", position="dodge", fill="#F8766D") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8)) + 
  scale_y_continuous("Med ABV (%)",label=scales::percent)  + 
  labs(title="Median ABV per State")


# comparing median IBU per state
summary_df %>% arrange(desc(Median_IBU)) %>% 
  mutate(State=fct_reorder(State, Median_IBU)) %>%
  ggplot(aes(State, Median_IBU,fill=State)) + 
  geom_bar(stat="identity", position="dodge",fill="#00BFC4") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8))+ 
  labs(title="Median IBU per State")

```


## Analysis #5
_We next examined which state has the most alcoholic (ABV) most bitter (IBU) beer. 
We found that Colorado has the most alcoholic beers while Oregon had the most bitter. _
```{r}
## comparing ABV vs IBU max

# myColors <- c("#7CAE00", "#C77CFF")
myColors <- c("#56ddc5", "#4699dd")
summary_df_pivot %>% filter(variable=="max_ABV.scaled" | variable=="max_IBU.scaled") %>% 
  ggplot(aes(State,value,fill=variable)) + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8)) + 
  geom_bar(stat="identity", position="dodge") + 
  scale_fill_manual(values=myColors) + 
  labs(y="Scale", title="Comparison of Max ABV vs IBU per State")

# comparing ABV per state
summary_df %>% arrange(desc(Max_ABV)) %>% mutate(State=fct_reorder(State, Max_ABV)) %>% slice(1:10) %>% 
  ggplot(aes(State, Max_ABV,fill=State)) + geom_bar(stat="identity", position="dodge",fill="#56ddc5") + theme(legend.position="none") + 
  scale_y_continuous("Max ABV (%)",label=scales::percent) + labs(title="Top 10 States with Most Alcoholic Beers")

# comparing IBU per state
summary_df %>% arrange(desc(Max_IBU)) %>% mutate(State=fct_reorder(State, Max_IBU)) %>% slice(1:10) %>%
  ggplot(aes(State, Max_IBU,fill=State)) + geom_bar(stat="identity", position="dodge",fill="#4699dd") + 
  ylab("Max IBU") + theme(legend.position="none") + labs(title="Top 10 States with Most Bitter Beers")
```


## Analysis #6
_Another aspect of the beer's ABV we looked into was the distribution and general statistics of ABV across all beers. 
Here we found that most of the beers fell within a normal distribution with most beers generally being somewhat equal in its 
alcoholic content. One thing to note in particular was how  _
```{r,warning=FALSE,message=FALSE}

summary(beer_df_cleaned$ABV)

summary_ABV <- beer_df_cleaned %>% ggplot(aes(beer_df_cleaned$ABV)) +  geom_histogram(aes(y=..density..), bindwidth=50)+
 geom_density()
summary_ABV


library(psych)
describe(beer_df_cleaned$ABV)
```


## Analysis #7 
_Relationship between IBU and ABV?_
```{r,warning=FALSE,message=FALSE}

# logging variables to see if it helps improve correlation
beer_df_cleaned$ABV.log <- log(beer_df_cleaned$ABV)
beer_df_cleaned$IBU.log <- log(beer_df_cleaned$IBU)

# install.packages("ggExtra")
library(ggExtra)

ABV_IBU_rel <- beer_df_cleaned %>% group_by(State) %>% ggplot(aes(x=ABV, y=IBU)) + geom_point() + geom_smooth(method="lm") +
  scale_x_continuous("Max ABV (%)",label=scales::percent) + 
  labs(title="Relationship between ABV and IBU",
       caption="Source: beer, brewery data")
ABV_IBU_rel_hist <- ggMarginal(ABV_IBU_rel, type="histogram", fill="slateblue")#, xparams=list(bins=10), yparams=list(bins=10))
ABV_IBU_rel_hist

cor(beer_df_cleaned$ABV, beer_df_cleaned$IBU)
cor(beer_df_cleaned$ABV.log, beer_df_cleaned$IBU.log)
cor(beer_df_cleaned$ABV, beer_df_cleaned$IBU.log)
cor(beer_df_cleaned$ABV.log, beer_df_cleaned$IBU)
```


## Analysis #8
_Difference of IBU and ABV between IPAs and other Ales?_
```{r,warning=FALSE,message=FALSE}
# goal here is to use kNN to classify if a beer is Ale/IPA/Other based on it's ABV and IBU
# viewing all types of beer styles
beer_df_cleaned %>% group_by(Style) %>% select(Style)

# # converting existing beers into type categories
# beer_df_cleaned$Style[grep("Ale", beer_df_cleaned$Style)]
# beer_df_cleaned$Style[grep("IPA", beer_df_cleaned$Style)]
# beer_df_cleaned$Style[grep("Lager", beer_df_cleaned$Style)]

# creating string match pattern
beer_df_cleaned$Beer_Category <- ifelse(grepl("Ale", beer_df_cleaned$Style), "Ale", 
                                        ifelse(grepl("IPA", beer_df_cleaned$Style) | grepl("India Pale Ales", beer_df_cleaned$Style), "IPA", "Other"))
head(beer_df_cleaned, 5)

beer_df_cleaned %>% group_by(Style) %>% select(IBU, Style) %>% summarise(Mean_IBU=mean(IBU)) %>% ggplot(aes(x=Style, y=Mean_IBU))+geom_point() + geom_smooth(method='lm')
head(beer_df_cleaned, 5)

# kNN
library(class) # for knn
library(caret) # for confusionMatrix
# match("ABV", names(beer_df_cleaned)) # column 7
# match("IBU", names(beer_df_cleaned)) # column 8
# running kNN k=5


# knn.cv(beer_df_cleaned[,c(7,8)],beer_df_cleaned$Beer_Category,prob = TRUE, k = 5)
beer_df_cleaned$pred_beer_cat <- knn.cv(beer_df_cleaned[,c(7,8)],beer_df_cleaned$Beer_Category,prob = TRUE, k = 5)

confusionMatrix(table(beer_df_cleaned$pred_beer_cat,beer_df_cleaned$Beer_Category))
```


## Analysis #9 
_EDA inference_
```{r}
# loading state summary data - population and elevation
state_data <- read.csv("C:/Git/DS6306_David_Neil_Project_1/DS6306_David_Neil_Project_1/Data Sources/Mean_Elevation_per_State.csv")
state_data <- state_data %>% plyr::rename(c("Mean.elevationList.by.state"="Mean_Elevation"))

# tidying up some of the data
state_data$Mean_Elevation <- gsub("feet", "", state_data$Mean_Elevation)
state_data$Mean_Elevation <- gsub(",", "", state_data$Mean_Elevation)
state_data$Mean_Elevation <- as.integer(state_data$Mean_Elevation)
state_data$Population <- gsub(",", "", state_data$Population)
state_data$Population <- as.integer(state_data$Population)


# adding elevation and population datra to summary_df
state_data <- state_data %>% select(State, State_Abr, Mean_Elevation, Population)
summary_df <- merge(x=summary_df, y=state_data, by.x="State", by.y="State_Abr", all.x=TRUE)


# counting distinct number of beers and breweries per state
state_beer_summary <- beer_df_cleaned %>% group_by(State) %>% summarise(Beers = n_distinct(Beer_ID))
state_brewery_summary <- beer_df_cleaned %>% group_by(State) %>% summarise(Breweries = n_distinct(Brewery_ID))


# adding distinct number of beers and breweries to summary_df
summary_df <- merge(x=summary_df, y=state_beer_summary, by.x="State", by.y="State", all.x=TRUE)
summary_df <- merge(x=summary_df, y=state_brewery_summary, by.x="State", by.y="State", all.x=TRUE)


# adding beers per million and breweries per million
summary_df <- summary_df %>% mutate(per_capita_beers = (Beers / (Population/1000000)))
summary_df <- summary_df %>% mutate(per_capita_breweries = (Breweries / (Population/1000000)))

summary_df <- summary_df %>% plyr::rename(c("State"="State_Abvr", "State.y"="State"))

# reording the columns
summary_df <- summary_df[c(12,1,2,3,4,5,6,7,8,9,10,11,13,14,15,17,16,18)]

library(maps)
us_states <- map_data("state")

summary_df$State_lwrcase <- tolower(summary_df$State)
head(state_data, 5)

us_states_data <- merge(x=us_states, y=summary_df, by.x="region", by.y="State_lwrcase")


p0 <- ggplot(data = us_states_data,mapping = aes(x = long, y = lat, group = group, fill = per_capita_breweries))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) + coord_map(projection = "albers", lat0 = 39, lat1 = 45) + labs(title = "x") + theme_map() + labs(fill = "Percent")
p1

p2 <- p1 + scale_fill_gradient(low = "white", high = "#CB454A") + labs(title = "Per Capita Breweries per 1M Drinking Age Adults") + theme_map() + labs(fill = "Percent")
p2

```
